---
title: "PASI2"
author: "Gerardo Fumagal"
date: "2/15/2021"
output: 
  html_document:
    df_print: paged
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of ToothGrowth data set

The ToothGrowth data set in r, is a collection of observations on guinea pigs
tooth growth due to the application of Vitamin C dosis. In this paper, we will
analyze the content of this data set, by getting information on the number of
observations, the variables captured and by doing basis statistical analysis of
the data with preliminary conclusions.

```{r Reading and storing data}
data = ToothGrowth
```

###Exploratory Data Analisys

To get an understanding of the ToothGrowth set, we will explore the data as per
the following steps:

1. Number of observations captured.
2. Number, name and type of variables within the data set.
3. Existing levels of variables, and unique values for each.

Let's explore the data set:

```{r Exploratory Data Analysis}
dim(data)
str(data)
```

ToothGrowth data set is a collection of 60 observations containing the variables
tooth length, supplementary type and dose. Length is a numeric variable, 
supplementary type is a factor variable with 2 levels and dose is another
numeric variable.

Now we get more information by making a summary on the database and extracting
the unique values from dose variable:

```{r Exploratory Data Analysis 2}
summary(data)
unique(data$dose)
```

We noticed there were 2 groups for testing each of 30 pigs identified by the 
supplementary type variable as VC and OC, they received either one of 3 
different doses level: 0.5, 1.0 or 2.0 units.

## Statistical Analysis

The main goal in this section is to find out relationships between the length of
tooth with dose and supplementary type. 

Let's start with supplementary type, a variable with 2 levels: VC or OJ. We need
to, find out the distribution for each level

```{r Boxplot Supplementary type vs Tooth length}
plot(x = data$supp, y = data$len, xlab = "Supplementary Type", ylab = "Tooth length")
```

By looking at the graph, there is a perception of OJ median is higher than the 
one for VC, so let this be our first hypothesis.

The null hypothesis is the difference between the expected value of the tooth 
length with supplementary type OJ is equal to the expected value with type VC,
the alternative is, the expected value with supplementary type OJ is higher than
with type VC:

\begin{equation}
H_{0} \textrm{ | }  \mu_{OJ} = \mu_{VC} \\
H_{a} \textrm{ | }  \mu_{OJ} > \mu_{VC}
\end{equation}

To test our hypothesis we consider that, the groups are the same size and 
independent, a t-test is to be used with an average of the variances as the pool
variance.

```{r t-test for Supplementary type vs Tooth length}
x_oj <- with(data, len[supp == "OJ"])
x_vc <- with(data, len[supp == "VC"])
t.test(x_oj - x_vc, paired = FALSE, var.equal = TRUE)
```

Based on the output from the t-test, we reject the null hypothesis as our 95%
confidence interval (1.408659 5.991341), does not include 0, and as the interval
is clearly on the positive side, we can conclude the supplementary type OJ has
a positive effect on tooth length.

## Analysis of dose and supplementary type effects on tooth length

Let's visualize the data by dose and supplementary type at the same time, in 
order to recognize some preliminary patterns.

```{r Plot Tooth length by dose and supplementary type}
plot(data$dose, data$len, col = data$supp, ylab = "Tooth length", xlab = "dose")
```


## Including Plots

You can also embed plots, for example:

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
