---
title: "Exponential distribution "
author: "Gerardo Fumagal"
date: "2/12/2021"
header-includes:
   - \usepackage{amsmath}
   - \usepackage{mathtools}
   - \newcommand{\en}[1]{{\selectlanguage{english}#1}}
   - \newcommand{\gr}[1]{{\selectlanguage{greek}#1}}
mainfont: GFS Didot
output:
  html_document:
    df_print: paged
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exponential distribution and CLT

The exponential distribution is the probability distribution of the time between
events in a Poisson point process in which events occur constantly and independently at
a constant average rate.

The PDF of an exponential distribution is:

\begin{equation}
f(x; \alpha) = 
\begin{cases}
\lambda e^\lambda & x \geq 0\\
0                 & x \lt 0.\\
\end{cases}
\end{equation}

Where $\lambda$ is the key parameter, the mean of the distribution and standard 
deviation are defined as follows:

\begin{equation}
\mu = \frac{\textrm{1}}{\lambda} ; \sigma = \frac{\textrm{1}}{\lambda}
\end{equation}


equal to both the mean and the standard deviation of the exponential
distribution.

In this paper, we will compare the exponential distribution with the central
limit theorem, by simulating the exponential distribution and illustrate its
properties.

CLT establishes that the distribution of averages of iid variables normalized,
becomes that of a standard normal, as the sample size increases.

If $X_{1} , X_{2} ... X_{i}$ are a collection of iid random variables with mean
$\mu$ and variance $\sigma_{2}$; also, let $\bar{X_{n}}$ be their sample average
then:

\begin{equation}
\frac{\bar{X_{n}} - \mu}{\sigma / \sqrt{n}}
\end{equation}

has a distribution of a that of a standard normal for large $n$.

It can be express as the following form:

\begin{equation}
\frac{\textrm{Estimate - Mean of estimate}}{\textrm{Std. error of estimate}}
\end{equation}

## Simulation

First, let's generate data for 1000 simulations of exponential distribution with
a sample size of 40, for the porpoise of this exercise will take $\lambda$ = 0.2.

We need to generate the exponential distribution sample, the means and variance,
and to normalize them according the the equation by CLT. The following code does
generate these simulation, calculate means and normalized them:

```{r simulate data}
lambda <- 0.2
expn <- 100
n <- 1000
meanr <- NULL
meanavg <- NULL
var_exp <- NULL
exp_tonormal <- NULL
var_avg <- NULL
for (i in 1:n) {
  meanr = c(meanr, mean(rexp(expn, lambda)))
  meanavg = c(meanavg, mean(meanr))
  var_exp = c(var_exp, var(rexp(expn, lambda)))
  var_avg = c(var_avg, mean(var_exp))
  SE <- sqrt(var(rexp(expn, lambda))/expn)
  exp_tonormal = c(exp_tonormal,(mean(rexp(expn, lambda))-(1/lambda))/SE)
}

```

## Comparing Exponential Distribution with CLT

1. Sample mean and theoretical mean

We discuss that exponential distribution has a mean of equal to 1/$\lambda$, in
our case this value is 5, let's plot how the simulation behaves as the mean of
average sample towards the expected value.

```{r sample mean and theoretical mean, echo=FALSE}
mean_theo <- 1/lambda
plot(meanavg, type = "l", main = "Exponential Distribution, sample mean vs 
     theoretical mean", ylab = "cumulative means" , xlab = "iteration")
abline(h=mean_theo, lwd = 2, col = "red")
```

The plot shows how the simulation converges to the expected value, our sample 
size of 40 seems to be large enough to drive the result to the expected value.

2. Compare how variable the sample is by simulating the variance vs theoretical
variance value.

Once again, the standard deviation of the exponential distribution is 1/$\lambda$
which it is also 5, which makes its variance ($S$ = $\sigma^2$) equal to 25.

Let's plot the results from our simulation and note how they behaves towards the 
theoretical variance value.

```{r, sample vairance vs theoretical variance}
var_theo <- (1/lambda)^2
plot(var_avg, type = "l", main = "Exponential Distribution, sample variance vs 
     theoretical variance", ylab = "cumulative variance" , xlab = "iteration")
abline(h=var_theo, lwd = 2, col = "red")

```

The simulation converges to the function's variance, the sample size is large
enough to show the variance value is 25.

3. Is the distribution approximately normal?

We have applied CLT formula on the averages of the samples and store in a 
variable called exp_tonormal, let's plot the results and compare to a normal 
distribution:

```{r, plot exponential distribution vs normal}
h <- hist(exp_tonormal, breaks=30, col="red", xlab="x", 
   main="Exponential Simulation vs CLT sample size = 40") 
xfit <- seq(min(exp_tonormal), max(exp_tonormal), length = 40)
yfit<-dnorm(xfit,mean=mean(exp_tonormal),sd=sd(exp_tonormal))
yfit <- yfit*diff(h$mids[1:2])*length(exp_tonormal)
lines(xfit, yfit, col = "blue", lwd = 2)
```


Now we can tell that the exponential distribution is approximately a normal with 
a sample size of 40, although, it may be more clear if we increase the size
just to see this effect, let's try a sample size of 1000, and plot the results.

```{r sample size 1000}
expn <- 1000
n <- 1000
exp_tonormal <- NULL
for (i in 1:n) {
  SE <- sqrt(var(rexp(expn, lambda))/expn)
  exp_tonormal = c(exp_tonormal,(mean(rexp(expn, lambda))-(1/lambda))/SE)
}
```

```{r, plot exponential distribution vs normal sample size 200, echo=FALSE}
h <- hist(exp_tonormal, breaks=30, col="red", xlab="x", 
   main="Exponential Simulation vs CLT sample size = 1000") 
xfit <- seq(min(exp_tonormal), max(exp_tonormal), length = 40)
yfit<-dnorm(xfit,mean=mean(exp_tonormal),sd=sd(exp_tonormal))
yfit <- yfit*diff(h$mids[1:2])*length(exp_tonormal)
lines(xfit, yfit, col = "blue", lwd = 2)
```

Now with a sample size of 200, the function fits better to a normal distribution
it is expected to be seamless as we increase the sample size
